---
title: Wild Boar are from Mars
subtitle: Do males move more than females?
author: Lizzy Nickell
output: html_document
---

```{r, echo = FALSE, warning=FALSE, message=FALSE}
#Housekeeping:

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

#Please load the following Libraries:
library(ComputationalMovementAnalysisData)
library(sf)
library(tidyverse)
library(psych)
library(ggpubr)
library(dplyr)
library(tmap)
```

## Abstract

There are many physical and behavioral differences between male and female wild boars. Females are typically smaller, and stay together in herds with other females and young. Males, on the other hand, wander larger territories, avoiding other males, and only interacting with females during mating season. We wanted to know if this difference in behavior is noticeable in their movement patterns. Using current data and multiple methods, we have found no differences in the movement patterns. While they may act differently, one sex does not move more than the other.

## Introduction

```{}
```

```{r}
# include plots automatically

ggplot(schreck_locations, aes(E, N)) +
  geom_point() +
  coord_map()

```

## Material and Methods

<!-- the following is just a placeholder text, remove it!-->

Christianity revaluation value battle faithful marvelous society derive free truth. Right battle mountains superiority grandeur ascetic grandeur merciful. Derive against intentions burying salvation ocean. Right intentions dead victorious ideal spirit evil ultimate joy. Holiest spirit value oneself contradict aversion christianity ultimate convictions war christian.

## Results

<!-- the following is just a placeholder text, remove it!-->

Philosophy oneself passion play fearful self noble zarathustra deceptions sexuality. Endless ocean of oneself dead ocean. Selfish decrepit.

## Discussion

<!-- the following is just a placeholder text, remove it!-->

Justice convictions spirit sexuality insofar free marvelous joy. Revaluation virtues mountains spirit fearful sexuality love endless. Society intentions will noble burying aversion moral. Insofar passion ultimate mountains of play gains depths joy christian reason christianity mountains dead. Mountains christianity play war holiest ascetic passion oneself derive grandeur. Against pinnacle hope joy burying ocean of horror disgust victorious faithful justice suicide.
