---
title: Wild boar are from mars
subtitle: Do male boars move more than females?
author: He Yelu and Lizzy Nickell
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

Proposal:

Formulate at least two research questions for your project-- Typical
types of research questions for a methodological project start with "How
can this and that be conceptualized/modelled/implemented...?" or ' "Can
this and that pattern be efficiently and effectively detected in this
and that data...?"

Possible Questions:

-   Is there a difference in how much males move vs. females?

Develop a research plan for investigating the above research
questions.--

-   What data do you need? Movement data from multiple individuals,
    equal number of males and females

-   Where do you get this data from? From the provided boar data set

-   What tools do you need? R, arcGIS (maybe)

-   Will you use extensions of tools you got to know in E1 -- E5?--

-   Are there other R tools or even tools beyond R that you will be
    using in your project?

Summarize your research plan in a semester project proposal.-- Use the
template given via github-- Submit your proposal until 15.05.2021

<!-- You can add  your R Code with Code chunks-->

```{r, echo = FALSE, warning=FALSE, message=FALSE}

# You can set chunk options individually per code chunk, as was done with this
# code chunk.

# echo = FALSE           hides the code from the generated output
# warning = FALSE        hides warnings from the generated output
# message = FALSE        hides messages from the generated output

# To set the setting for all code chunks, use the following code:

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

```

## Abstract

Wild boars show a distinct physical sexual dimorphism, with males being
much larger than the females. Can this difference be seen in other
aspects of wild boars? We analyzed the movement patterns of XX boars, X
male and X females to compare the amount of movement in the sexes. We
predicted that females move less than males, as they must care for the
young, and are not responsible for maintaining territory.

## Introduction

The wildboar (*Sus scrofa*) is native to most of Eurasia, and are
considered a pest in most countries. Because of this, numerous tracking
programs have been put in place to monitor their activities.

```{r}
library("devtools")
devtools::install_github("ComputationalMovementAnalysis/ComputationalMovementAnalysisData")
library(ComputationalMovementAnalysisData)
library(ggplot2)
library(tidyverse)
head(wildschwein_BE)

##   TierID TierName CollarID         DatetimeUTC       E       N
## 1      1     Ueli    12272 2014-05-28 21:01:14 2570390 1204820
## 2      1     Ueli    12272 2014-05-28 21:15:18 2570389 1204826
## 3      1     Ueli    12272 2014-05-28 21:30:13 2570391 1204821
## 4      1     Ueli    12272 2014-05-28 21:45:11 2570388 1204826
## 5      1     Ueli    12272 2014-05-28 22:00:33 2570388 1204819
## 6      1     Ueli    12272 2014-05-28 22:15:16 2570384 1204828
# Include tables with the function "kable"

knitr::kable(head(wildschwein_BE))
```

```{r}
# include plots automatically
ggplot(schreck_locations, aes(lat, lon)) +
  geom_point(aes(color = id)) 

ggplot(wildschwein_BE, aes(E,N))+
  geom_point(aes(color=TierID))
```

## Material and Methods

<!-- the following is just a placeholder text, remove it!-->

1) Clean data

2) combine data to assign sexes to boar id.

3) Assign scores to movement

4) Compare scores of males and females.

## Results

<!-- the following is just a placeholder text, remove it!-->

Philosophy oneself passion play fearful self noble zarathustra
deceptions sexuality. Endless ocean of oneself dead ocean. Selfish
decrepit.

## Discussion

<!-- the following is just a placeholder text, remove it!-->

Justice convictions spirit sexuality insofar free marvelous joy.
Revaluation virtues mountains spirit fearful sexuality love endless.
Society intentions will noble burying aversion moral. Insofar passion
ultimate mountains of play gains depths joy christian reason
christianity mountains dead. Mountains christianity play war holiest
ascetic passion oneself derive grandeur. Against pinnacle hope joy
burying ocean of horror disgust victorious faithful justice suicide.
