---
title: "Semester Project Review"
author: "Lizzy Nickell"
date: '2022-09-08'
output:
  html_document: default
  pdf_document: default
---

---
title: Wild Boar are from Mars
subtitle: Do males move more than females?
author: Lizzy Nickell
output: html_document
---

```{r Chunk 1, message=FALSE, warning=FALSE, include=FALSE}
#Housekeeping:

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

#Please load the following Libraries:
library(ComputationalMovementAnalysisData)
library(sf)
library(tidyverse)
library(psych)
library(ggpubr)
library(dplyr)
library(tmap)

```

## Abstract

There are many physical and behavioral differences between male and female wild boars. Females are typically smaller, and stay together in herds with other females and young. Males, on the other hand, wander larger territories, avoiding other males, and only interacting with females during mating season. These behavioral patterns can lead to the development of different deterrence methods. Using current data and multiple methods, there is some evidence that the sample population follows expected patterns. This evidence can be used to create sex-specific detterence methods.

## Introduction

The wild boar (Sus scrofa) is a highly adaptable species native to Europe and Asia and has been introduced to North America. They are highly invasive in North America, and considered pests in most of Europe and Asia, causing damage to crops and automobile damage due to vehicle impacts. As such, numerous projects have been put in place to track the movements of wild boar and monitor efforts to repel them from croplands and roads.

Wild boars are a dimorphic species. Males are 20% larger than females, weighing between 75 and 100kg, and 75-80cm at shoulder height (Marsan and Mattioli, 2013). Males are solitary for the majority of the year, seeking out females only during breeding season (Wickline, 2014). females are smaller, weighing between 60 and 80kg, and only 63-65cm at shoulder height (Marsan and Mattioli, 2013). Females travel in herds anywhere from 6 to 100 individuals with piglets, juveniles, and males up to 2 years old (Wickline, 2014). Females protect offspring, sometimes leaving piglets with a single female while others go off to forage. Because of this difference in behavior, the typical home range also varies by sex. It has been estimated that home ranges for females are 500-1,000 ha, while males have a home range of 1,000-2,000 ha (Kueling and Leus, 2019).

One project enacted to monitor wild boar movement was conducted by WILMA and ZHAW in 2014 to 2016. Nineteen wild boar were tagged and monitored, with an effort to tag as many different herds as possible (Suter, 2013). Six of the boars were adult males, and the rest were adult females. Two main site locations were determined by looking at the center of the the boar's locations: Bern and Aargau. Some of the boars were captured multiple times and given new collars, but their ID was kept the same. Depending on the boar, and time of year, locations were taken at different sampling ratios. Boars with enough data were used for this analysis.

Analyzing the differences in movement based on sex is an important step for evaluating deterrence methods, as different movement patterns could be more reactive to different methods. It is expected that males will have larger areas, but travel alone, so wider-spread deterrence methods, like acoustic-scare devices, would be more effective. Females likely travel in large herds, but smaller areas, implying that a more concentrated deterrence, like electric fences, would be more effective. Confirming that the boars in the study follow the expected pattern is the first step in establishing effective deterrence strategies.

```{r Chunk 2, include=FALSE}
#load data
pigs <- wildschwein_BE
metapigs <- wildschwein_metadata

#add sex to base data
sexypigs <- read.csv("~/Documents/cma_project/SemesterProject_Lizzy/sexypigs.csv")

sexypigs <- st_as_sf(sexypigs, coords = c("E", "N"), 
                          crs = 2056,
                          remove = FALSE)

sexypig1 <- filter(sexypigs, TierID == "1")

tmap_mode("view")

map0 <- tm_shape(sexypig1) +
  tm_dots(col = "green")
map0
```

## Material and Methods

Nineteen wild boar were tagged and monitored between the years of 2014 and 2016. Six of the boars were adult males, and thirteen were adult females. There were two main sites that the boars were found in: Bern and Aargau. Locations were determined by looking at the center of the the boar's locations. Depending on the boar, and time of year, locations were taken at different sampling ratios. Over the time period, enough location points were collected to get an idea of the area covered by each boar. The data was compiled into an R package and into two data frames, the main data, and a meta data set. The data was then converted from a standard data frame to a sf object.

Three methods were used to analyze the motion of the pigs and compare them by sex. The first was a convex hull calculation, calculating the total area of land covered by the wild boars. A convex hull is the smallest possible polygon that can be created containing all of the points of interest. Figure 1 shows the convex hull overlaying the location points of boar number 1, Ueli.

First, the data was filtered by sex, and using the sf package, the filtered data was grouped, summarized, and converted into a convex hull. The area of each polygon was then calculated. The two groups were recombined, and statistical tests were run to compare the areas by sex. Shapiro tests showed that the data was not normal, so a wilcox test was run to test significance.

```{r include=FALSE}
#convex hull

devtools::install_github("tidyverse/dplyr")


malepigs <- filter(sexypigs, sex == "m")
femalepigs <- filter(sexypigs, sex == "f")

mpigsgrouped <- group_by(malepigs,TierID)
mpigssmry <- summarise(mpigsgrouped)
polymales <- st_convex_hull(mpigssmry)
mareas <- polymales %>% 
  mutate(area = st_area(polymales))

fpigsgrouped <- group_by(femalepigs,TierID)
fpigssmry <- summarise(fpigsgrouped)
polyfemales <- st_convex_hull(fpigssmry)
fareas <- polyfemales %>%
  mutate(area = st_area(polyfemales))

convexpigs <- read.csv("~/Documents/cma_project/SemesterProject_Lizzy/convexpigs.csv")

group_by(convexpigs, Sex) %>%
  summarise(
    count = n(),
    mean = mean(ConvexArea, na.rm = TRUE),
    sd = sd(ConvexArea, na.rm = TRUE))

with(convexpigs, shapiro.test(ConvexArea[Sex == "m"]))
with(convexpigs, shapiro.test(ConvexArea[Sex == "f"]))
#not normal
wilcox.test(ConvexArea ~ Sex, data = convexpigs,
            exact = FALSE)
#significant
```

```{r echo=FALSE}
pig01 <- polymales%>% filter(TierID == 1)

map1 <- tm_shape(sexypig1) +
  tm_dots(col = "green") + 
  tm_shape(pig01) +
  tm_polygons(col = "#00cef6",alpha = 0.4,border.col = "#00cef6")

map1
```

*Figure 1: Convex Hull of wild boar 1, "Ueli."*

The second method was a traditional bounding box method, taking the extreme coordinates in each direction and creating a box from those coordinates. Figure 2 shows the resulting bounding box from boar one's data overlaying the convex hulls and location data. Each line of the box is the created taking the furthest points and creating lines parallel to longitude and latitude. The area of each box was calculated, then the averages for each sex were calculated. Like the convex hull averages, the data was not normal, and a wilcox test was run to test significance.

```{r include=FALSE}
#Bounding Boxes
mcp <- sexypigs
mcp <- st_as_sf(mcp, coords = c("E", "N"), 
                          crs = 2056,
                          remove = FALSE)
#Isolate piggies
pig1 <- mcp %>% filter(TierID == 1)
pig2 <- mcp %>% filter(TierID == 2)
pig5 <- mcp %>% filter(TierID == 5)
pig10 <- mcp %>% filter(TierID == 10)
pig11 <- mcp %>% filter(TierID == 11)
pig16 <- mcp %>% filter(TierID == 16)
pig18 <- mcp %>% filter(TierID == 18)
pig22 <- mcp %>% filter(TierID == 22)
pig25 <- mcp %>% filter(TierID == 25)
pig31 <- mcp %>% filter(TierID == 31)
pig36 <- mcp %>% filter(TierID == 36)
pig40 <- mcp %>% filter(TierID == 40)
pig48 <- mcp %>% filter(TierID == 48)
pig60 <- mcp %>% filter(TierID == 60)
pig83 <- mcp %>% filter(TierID == 83)
pig84 <- mcp %>% filter(TierID == 84)
pig86 <- mcp %>% filter(TierID == 86)
pig97 <- mcp %>% filter(TierID == 97)
pig98 <- mcp %>% filter(TierID == 98)
  
  #Calculate bounding boxes
box1 <- st_bbox(pig1)
box2 <- st_bbox(pig2)
box5 <- st_bbox(pig5)
box10 <- st_bbox(pig10)
box11 <- st_bbox(pig11)
box16 <- st_bbox(pig16)
box18 <- st_bbox(pig18)
box22 <- st_bbox(pig22)
box25 <- st_bbox(pig25)
box31 <- st_bbox(pig31)
box36 <- st_bbox(pig36)
box40 <- st_bbox(pig40) 
box48 <- st_bbox(pig48)
box60 <- st_bbox(pig60)
box83 <- st_bbox(pig83)
box84 <- st_bbox(pig84)
box86 <- st_bbox(pig86)
box97 <- st_bbox(pig97)
box98 <- st_bbox(pig98)

  #calculate area of bounding boxes
area1 <- ((box1$ymax-box1$ymin) * (box1$xmax-box1$xmin))
area2 <- ((box2$ymax-box2$ymin) * (box2$xmax-box2$xmin))
area5 <- ((box5$ymax-box5$ymin) * (box5$xmax-box5$xmin))
area10 <- ((box10$ymax-box10$ymin) * (box10$xmax-box10$xmin))
area11 <- ((box11$ymax-box11$ymin) * (box11$xmax-box11$xmin))
area16 <- ((box16$ymax-box16$ymin) * (box16$xmax-box16$xmin))
area18 <- ((box18$ymax-box18$ymin) * (box18$xmax-box18$xmin))
area22 <- ((box22$ymax-box22$ymin) * (box22$xmax-box22$xmin))
area25 <- ((box25$ymax-box25$ymin) * (box25$xmax-box25$xmin))
area31 <- ((box31$ymax-box31$ymin) * (box31$xmax-box31$xmin))
area36 <- ((box36$ymax-box36$ymin) * (box36$xmax-box36$xmin))
area40 <- ((box40$ymax-box40$ymin) * (box40$xmax-box40$xmin))
area48 <- ((box48$ymax-box48$ymin) * (box48$xmax-box48$xmin))
area60 <- ((box60$ymax-box60$ymin) * (box60$xmax-box60$xmin))
area83 <- ((box83$ymax-box83$ymin) * (box83$xmax-box83$xmin))
area84 <- ((box84$ymax-box84$ymin) * (box84$xmax-box84$xmin))
area86 <- ((box86$ymax-box86$ymin) * (box86$xmax-box86$xmin))
area97 <- ((box97$ymax-box97$ymin) * (box97$xmax-box97$xmin))
area98 <- ((box98$ymax-box98$ymin) * (box98$xmax-box98$xmin))

  #Regroup into data frame
areas <- c(area1, area2, area5, area10, area11, area16, area18, area22, area25, area31, area36, area40, area48, area60, area83, area84, area86, area97, area98 )
namareas <- c("area1", "area2", "area5", "area10", "area11", "area16", "area18", "area22", "area25", "area31", "area36", "area40", "area48", "area60", "area83", "area84", "area86", "area97", "area98")

df <- data.frame(namareas, areas)
df

write.csv(df, file = "boxypigs.csv")
#add sex
boxypigs1 <- read_csv("boxypigs1.csv")


  #Calculate differences by sex
group_by(boxypigs1, Sex) %>%
  summarise(
    count = n(),
    mean = mean(areas, na.rm = TRUE),
    sd = sd(areas, na.rm = TRUE)
  )

with(boxypigs1, shapiro.test(BoxArea[Sex == "m"]))
with(boxypigs1, shapiro.test(BoxArea[Sex == "f"]))
#not normal
wilcox.test(BoxArea ~ Sex, data = boxypigs1,
            exact = FALSE)

#not significant
```

```{r echo=FALSE}

#Pig1 example
pig1 <- mcp %>% filter(TierID == 1)
  
box1 <- st_bbox(pig1)
box1 = st_as_sfc(box1)
box1 <- st_as_sf(box1, coords = c("E", "N"), 
                          crs = 2056,
                          remove = FALSE)

map2 <- tm_shape(sexypig1) +
  tm_dots(col = "green") +
  tm_shape(pig01) +
  tm_polygons(col = "#00cef6",alpha = 0.4,border.col = "#00cef6") +
  tm_shape(box1)+
  tm_polygons(alpha = 0,border.col = "Black")
map2
```

*Figure 2: Example of a bounding box created from wild boar number 1 overlaid with the convex polygon created.*

The third method was an adaptation of the bounding box. The furthest coordinates in each direction were identified and a polygon was created connecting those four points, rather than the bounding box that created parallel lines (fig 3). The area was calculated for each pig, and the areas were averaged for each sex. A wilcox test was run to test for significant difference.

```{r include=TRUE, echo=FALSE}
#Min-max boxes

#sort data: min/max E & N
pigs2 <- read.csv("~/Documents/cma_project/SemesterProject_Lizzy/pigs2.csv")

  #Make data sf
pigs2 <- st_as_sf(pigs2, 
                          coords = c("E", "N"), 
                          crs = 2056,
                          remove = FALSE)

pigs2grouped <- group_by(pigs2,TierID)
pigs2smry <- summarise(pigs2grouped)

mmpigs <- st_convex_hull(pigs2smry)


pig1.1 <- mmpigs %>% filter(TierID == 1)

map3 <- tm_shape(sexypig1) +
  tm_dots(col = "green") +
  tm_shape(pig01) +
  tm_polygons(col = "#00cef6",alpha = 0.4,border.col = "#00cef6") +
  tm_shape(box1)+
  tm_polygons(alpha = 0,border.col = "Black")+
  tm_shape(pig1.1)+
  tm_polygons(alpha = 0.2, border.col= "Purple")
  
map3
```

*Figure 3: Min-max polygon created from wild boar 1's data (purple), overlaying the bounding box (black), and the convex polygon (blue).*

```{r Chunk 7: Min-Max Polygons, include=FALSE}

polypigs2 <- pigs2smry %>% 
  st_cast("POLYGON")
  
mmareapigs2 <-  polypigs2 %>%
  mutate(area = st_area(geometry))

write.csv(mmareapigs2, file = "pigareas.csv")

#add sex
mmpigs <- read.csv("~/Documents/cma_project/SemesterProject_Lizzy/mmpigs.csv")

group_by(mmpigs, Sex) %>%
  summarise(
    count = n(),
    mean = mean(mmArea, na.rm = TRUE),
    sd = sd(mmArea, na.rm = TRUE)
  )

with(mmpigs, shapiro.test(mmArea[Sex == "m"]))
with(mmpigs, shapiro.test(mmArea[Sex == "f"]))
  #not normal

wilcox.test(mmArea ~ Sex, data = mmpigs,
            exact = FALSE)

```

## Results

```{r include=FALSE}
conresults <- group_by(convexpigs, Sex) %>%
  summarise(
    count = n(),
    mean = mean(ConvexArea, na.rm = TRUE),
    sd = sd(ConvexArea, na.rm = TRUE))

boxresults <- group_by(boxypigs1, Sex) %>%
  summarise(
    count = n(),
    mean = mean(BoxArea, na.rm = TRUE),
    sd = sd(BoxArea, na.rm = TRUE))

mmresults <- group_by(mmpigs, Sex) %>%
  summarise(
    count = n(),
    mean = mean(mmArea, na.rm = F),
    sd = sd(mmArea, na.rm = F))

wilcox.test(ConvexArea ~ Sex, data = convexpigs,
            exact = FALSE)
wilcox.test(BoxArea ~ Sex, data = boxypigs1,
            exact = FALSE)
wilcox.test(mmArea ~ Sex, data = mmpigs,
            exact = FALSE)

conresults
boxresults
mmresults

```

```{r include=FALSE}
#figure 3
allboars <- print(list(convexpigs, boxypigs1, mmpigs)) %>% reduce(full_join, by = "TierID")
write_csv(allboars, file = "allboars.csv")

allboars <- read.csv("~/Documents/cma_project/SemesterProject_Lizzy/allboars2.csv")
```

```{r echo=FALSE}
boxplot<- ggplot(allboars, aes(x=Sex, y=Area, fill=Sex)) + 
  geom_boxplot()+
  facet_wrap(~Method)+
  labs(y= "Area (m^2)")
boxplot
```

*Figure 4: Box plots of the resulting areas from the Convex Hull (A), Bounding Box Method (B) and the Max-Min Polygon Method (C).*

Every method returned the expected results of males covering larger areas than females. The convex hull method had slightly significantly different averages (p=0.048), with males having an average range of 1713ha, and females having an average range of 557ha. Neither the bounding box method nor the min-max method were significantly different (0.12 and 0.21, respectively). As seen in figure 4, the bounding box method calculated larger areas than the convex method while the min-max method calculated smaller areas than the convex method. The ratio of female to male, however, were all similar, ranging from 0.26 to 0.33.

## Discussion

The convex hull calculations resulted in the the most significant results of the three different methods. As the most traditional and direct method, the convex hull is definitely method that best represents the area covered by individual boars. Through this method, it can be confirmed that the male boars in this study do cover more area than the the female boars. The next steps would be to test different densities of deterrence devices in areas that the different sexes frequent. Areas that are more commonly covered by males theoretically need less devices, or different devices, like acoustic vs electric, and further testing can prove that.

The convex hull calculations can be computationally taxing. To reduce the amount of calculations, the best established method is the bounding box method. As already mentioned, the bounding box methods result in higher areas than the convex hull method, and artificially inflate the results. Additionally the bounding box method loses any shape details. Looking again at figure 3, there are areas of land that that are covered by the bounding box that boar 1 has never visited. If the bounding boxes were used to determine placement of deterrence devices, those devices could be completely useless. Comparatively, this new min-max method underestimates the area covered, but ensures that all of the areas covered are actually areas visited by wild boar herds. A researcher can be assured that deterrence devices placed within this polygon will have a high chance of interacting with a boar.

It should be pointed out that the number of boars sampled in this study is very small. Small sample size reduces the power of these tests, and the results of the study, while confirming, may still have errors, and if more decisions are made based on this limited data, they could be the wrong decisions to make. When and if more data is collected, a less taxing computation will be useful. When comparing the bounding box and the min-max method, the min-max method will lead to more effective suggestions when it comes to where deterrence devices should be placed.

## References

Keuling, O., & Leus, K. (2019). Sus scrofa (Wild Boar). Retrieved June 30, 2022, from <https://www.iucnredlist.org/species/41775/44141833#habitat-ecology>

Marsan, Andrea., & Mattioli, Stefano. (2013). Il cinghiale.ISBN 978-88-96348-178.

Suter, S. (2013). Prevention of wild boar damage in agriculture. In WILMA/ZHAW.

Wickline, K. (2014). Sus scrofa. Retrieved June 3, 2022, from <https://animaldiversity.org/site/accounts/information/Sus_scrofa.html#880B0B11-4354-11E3-8D47-002500F14F28>

## Appendix A

Convex Hull Polygons of all Boars

```{r}

sexypigsgrouped <- group_by(sexypigs,TierID)
sexypigssmry <- summarise(sexypigsgrouped)
sexypoly <- st_convex_hull(sexypigssmry)

tmap_mode("view")

tm_shape(sexypoly) +
  tm_polygons(col = "#00cef6",alpha = 0.4,border.col = "#00cef6")+
  tm_facets(by= "TierID")
```
